Homework Report

### 2017-02-14 work/work.tex
Adding refs for heuristic lettering and updating crown graph definition

@@

\newtheorem{prop}{Proposition}
\newtheorem*{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem*{definition}{Definition}

@@

\subsection*{Heuristic A}

Heuristic A is the greedy algorithm.

@@

\subsection*{Heuristic B}

Heuristic B is the greedy algorithm with degree sequencing. It orders the vertices according to the decreasing value of their degree. This is also known as the Welsh-Powell algorithm, which is defined in \cite{welsh}.

@@

\subsection*{Heuristic C}

Heuristic C colors a graph by finding maximal independent sets of \(G\).

@@

\begin{prop}
Heuristic A and Heuristic B do not always produce an optimal solution, i.e. they do not always produce a minimum coloring of a graph.
\end{prop}

I will show this for \emph{heuristic B}. It is trivial to show the same for \emph{heuristic A} as \emph{heuristic B} includes the same steps as \emph{heuristic A}. I will construct a simple graph, \(G = (V, E)\), such that:

@@

Let \(\chi\) refer to the coloring of \(G\) generated by \emph{heuristic B}. Let \(\chi^{*}\) be the optimal coloring of \(G\).

@@

By applying \emph{heuristic B}, we get the following coloring. The numbers indicate the ordering of vertices before applying the heuristic. Any vertex of the same degree got assigned arbitrarily. This results in \(\chi(G) = 3\).

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{images/graph-2.png}
\caption{Coloring from applying heuristic B}
\end{figure}

@@

\begin{prop}
Heuristic C also does not always produce an optimal solution, i.e. it does not always produce a minimum coloring of a graph.
\end{prop}

@@

By applying heuristic C and finding maximal independent sets, we see that \(K\) is \emph{5-colorable}.

\begin{figure}[H]
\centering
\includegraphics[scale=0.38]{images/mis-2.png}
\caption{Coloring of \(K\) after applying heuristic C}
\end{figure}

This is a non-optimal solution. We can see that the subgraph created by vertices \(\{5, 6, 7, 8\}\) is complete and thus we need at least 4 colors. This graph is \emph{4-colorable} however, meaning \(\chi^{*}(K) = 4\). Thus, we can see that heuristic C does not always give us an optimal solution.

@@

\begin{prop}
Heuristic A and heuristic B have an upper bound of \(\Delta + 1\).
\end{prop}

@@

\begin{definition}[Crown Graph]
A crown graph \(CR_n = (V, E)\) is an undirected graph with two sets of vertices where \(V = V_1 \cup V_2\) with an edge from \(v_i \in V_1\) to \(v_{j} \in V_2\) whenever \(i \neq j\). A crown graph can also be viewed as a complete bipartite graph where the edges of a perfect matching have been removed.
\end{definition}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.38]{images/graph-5.png}
\caption{Worst-case coloring of \(H\) using either heuristic A or B}
\end{figure}

@@

The proof of \emph{Brook's Theorem} can be found in \cite{lovasz}. Overall, heuristic A and heuristic B can produce some very undesirable results. In graph \(H\), at the worst case, these heuristics produce \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite. This is shown below.

@@

\begin{prop}
Heuristic A and Heuristic B have a running time of $O(|V| + |E|)$.
\end{prop}

It is possible to implement heuristic A and heuristic B efficiently and achieve a linear running time \cite{kubale}. By using an adjacency list representation of a graph \(G = (V, E)\), the greedy algorithm will look once at each vertex and each edge twice. This is because for each vertex, all of its neighbors are checked. This leads to the algorithm crossing each edge twuce. As this is still linear, we know \(O(|V| + 2|E|) = O(|V| + |E|)\).

\begin{prop}
Heuristic A and Heuristic B produce a feasible, proper coloring of input graph \(G = (V, E)\).
\end{prop}

>>>>

\newcounter{heuristic} \setcounter{heuristic}{0}
\renewcommand{\theheuristic}{\Alph{heuristic}}

\newenvironment{heuristic}[1]%
{
\refstepcounter{heuristic}
\subsection*{Heuristic \theheuristic}
}%
{}

@@

\newtheorem{prop}{Proposition}
\newtheorem{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

@@

\heuristic{}\label{heuristic-greedy}

Heuristic \ref{heuristic-greedy} is the greedy algorithm.

@@

\heuristic{}\label{heuristic-welsh}

Heuristic \ref{heuristic-welsh} is the greedy algorithm with degree sequencing. It orders the vertices according to the decreasing value of their degree. This is also known as the Welsh-Powell algorithm, which is defined in \cite{welsh}.

@@

\heuristic{}\label{heuristic-mis}

Heuristic \ref{heuristic-mis} colors a graph by finding maximal independent sets of \(G\).

@@

\begin{prop}
Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} do not always produce an optimal solution, i.e. they do not always produce a minimum coloring of a graph.
\end{prop}

I will show this for \emph{heuristic \ref{heuristic-welsh}}. It is trivial to show the same for \emph{heuristic \ref{heuristic-greedy}} as \emph{heuristic \ref{heuristic-welsh}} includes the same steps as \emph{heuristic \ref{heuristic-greedy}}. I will construct a simple graph, \(G = (V, E)\), such that:

@@

Let \(\chi\) refer to the coloring of \(G\) generated by \emph{heuristic \ref{heuristic-welsh}}. Let \(\chi^{*}\) be the optimal coloring of \(G\).

@@

By applying \emph{heuristic \ref{heuristic-welsh}}, we get the following coloring. The numbers indicate the ordering of vertices before applying the heuristic. Any vertex of the same degree got assigned arbitrarily. This results in \(\chi(G) = 3\).

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{images/graph-2.png}
\caption{Coloring from applying heuristic \ref{heuristic-welsh}}
\end{figure}

@@

\begin{prop}
Heuristic \ref{heuristic-mis} also does not always produce an optimal solution, i.e. it does not always produce a minimum coloring of a graph.
\end{prop}

@@

By applying heuristic \ref{heuristic-mis} and finding maximal independent sets, we see that \(K\) is \emph{5-colorable}.

\begin{figure}[H]
\centering
\includegraphics[scale=0.38]{images/mis-2.png}
\caption{Coloring of \(K\) after applying heuristic \ref{heuristic-mis}}
\end{figure}

This is a non-optimal solution. We can see that the subgraph created by vertices \(\{5, 6, 7, 8\}\) is complete and thus we need at least 4 colors. This graph is \emph{4-colorable} however, meaning \(\chi^{*}(K) = 4\). Thus, we can see that heuristic \ref{heuristic-mis} does not always give us an optimal solution.

@@

\begin{prop}
Heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} have an upper bound of \(\Delta + 1\).
\end{prop}

@@

\begin{definition}[Crown Graph]
A crown graph \(CR_n = (V, E)\) is an undirected graph such that \(U = \{u_1, u_2, \dots, u_i\}\) and \(W = \{w_1, w_2, \dots, w_j\}\) where \(V = U \cup W\), \(U \cap W = \emptyset\), and \(|U| = |W|\). There is an edge from \(u_i\) to \(w_j\) whenever \(i \neq j\). A crown graph can also be viewed as a complete bipartite graph where the edges of a perfect matching have been removed.
\end{definition}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.38]{images/graph-5.png}
\caption{Worst-case coloring of \(H\) using either heuristic \ref{heuristic-greedy} or \ref{heuristic-welsh}}
\end{figure}

@@

The proof of \emph{Brook's Theorem} can be found in \cite{lovasz}. Overall, heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} can produce some very undesirable results. In graph \(H\), at the worst case, these heuristics produce \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite. This is shown below.

@@

\begin{prop}
Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} have a running time of $O(|V| + |E|)$.
\end{prop}

It is possible to implement heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} efficiently and achieve a linear running time \cite{kubale}. By using an adjacency list representation of a graph \(G = (V, E)\), the greedy algorithm will look once at each vertex and each edge twice. This is because for each vertex, all of its neighbors are checked. This leads to the algorithm crossing each edge twuce. As this is still linear, we know \(O(|V| + 2|E|) = O(|V| + |E|)\).

\begin{prop}
Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} produce a feasible, proper coloring of input graph \(G = (V, E)\).
\end{prop}
